---
title: "CLT: Sampling Distribution of the Mean, Median & Mode"
subtitle: "R simulation (knitr engine)"
format:
  html:
    page-layout: full
    code-fold: true
    code-summary: "Show code"
params:
  alpha: 0
  n: 2000
  reps: 3000
  seed: 42
---

```{r setup}
#| include: false
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# sn: skew-normal distribution (dsn, rsn, qsn)
# moments: skewness()
if (!requireNamespace("sn", quietly = TRUE)) {
  install.packages("sn")
}
if (!requireNamespace("moments", quietly = TRUE)) {
  install.packages("moments")
}
if (!requireNamespace("ggplot2", quietly = TRUE)) {
  install.packages("ggplot2")
}

library(sn)
library(moments)
library(ggplot2)
```


## 1. Pick a population distribution

Choose a skew for the parent population (using a skew-normal distribution),
then look at its shape. Edit `params$alpha` (or pass `-P alpha:VALUE` to
`quarto render`) to change the skew — no interactive widget is needed since
this is a static R render, but you can re-render with different parameter
values.

```{r population}
# params:
alpha <- 20
n <- 2000
reps <- 120000
seed <- 42

# skew-normal(xi = location, omega = scale, alpha = shape)
pop_x <- seq(
  sn::qsn(0.001, xi = 0, omega = 1, alpha = alpha),
  sn::qsn(0.999, xi = 0, omega = 1, alpha = alpha),
  length.out = 500
)
pop_pdf <- sn::dsn(pop_x, xi = 0, omega = 1, alpha = alpha)

pop_df <- data.frame(x = pop_x, density = pop_pdf)

ggplot(pop_df, aes(x = x, y = density)) +
  geom_line(color = "#3b6ea5") +
  geom_area(fill = "#3b6ea5", alpha = 0.2) +
  labs(
    title = paste0("Population distribution (skew \u03B1 = ", alpha, ")"),
    x = "x",
    y = "density"
  ) +
  theme_minimal()
```

## 2. Draw samples

For each repetition, draw a sample of size `n` from the population above.
Both `n` and the number of repetitions (`reps`) are set as document
parameters, so you can push these much higher than a marimo widget could
comfortably handle — this chunk is plain R, so it scales with your available
memory rather than a browser-based reactive runtime.

```{r draw-samples}

set.seed(seed)

# reps x n matrix of draws — this is the part that needed to scale.
# rsn() is vectorised; for very large reps*n consider batching if memory
# becomes an issue (e.g. loop over chunks of reps and rbind/aggregate).
samples <- matrix(
  sn::rsn(reps * n, xi = 0, omega = 1, alpha = alpha),
  nrow = reps,
  ncol = n
)

cat(sprintf(
  "Drew %d samples of size n = %d from the population above.\n",
  reps,
  n
))
```

## 3. Distribution of the mean, median and mode

For every sample, compute its mean, median, and (binned) mode, then look at
how those three statistics are themselves distributed.

```{r compute-statistics}
sample_mode <- function(row, bins = 60) {
  h <- hist(row, breaks = bins, plot = FALSE)
  i <- which.max(h$counts)
  (h$breaks[i] + h$breaks[i + 1]) / 2
}

sample_means <- rowMeans(samples)
sample_medians <- apply(samples, 1, median)
sample_modes <- apply(samples, 1, sample_mode)
```

```{r plot-statistics}
#| fig-width: 7
#| fig-height: 4
stat_df <- data.frame(
  value = c(sample_means, sample_medians, sample_modes),
  statistic = rep(c("mean", "median", "mode"), each = reps)
)

ggplot(stat_df, aes(x = value, fill = statistic)) +
  geom_density(alpha = 0.5) +
  scale_fill_manual(
    values = c(mean = "#3b6ea5", median = "#e08214", mode = "#5aa469")
  ) +
  labs(
    title = paste0("Sampling distributions (n = ", n, ", reps = ", reps, ")"),
    x = "statistic value",
    y = "density"
  ) +
  theme_minimal()
```

```{r summary-table}
summary_rows <- lapply(
  list(mean = sample_means, median = sample_medians, mode = sample_modes),
  function(v) {
    data.frame(
      avg = round(mean(v), 4),
      std_error = round(sd(v), 4),
      skewness = round(moments::skewness(v), 4)
    )
  }
)

summary_df <- do.call(rbind, summary_rows)
summary_df <- cbind(statistic = rownames(summary_df), summary_df)
rownames(summary_df) <- NULL

knitr::kable(
  summary_df,
  align = "c",
  caption = "Summary of sampling distributions"
)
```
